---
title: "ashishsa_hw4_p1"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(network)
library(igraphdata)
library(igraph)
library(GGally)
library(ape)
library(ggraph)
library(tidyverse)
```
##KARATE DATASET

We begin with making a network for the karate.We fit the karate dataset to the Heirarchical Random Graph.
```{r}
data(karate)
ghrg <- fit_hrg(karate)
```

We plot the Karate Dataset on a Graph and check the layout. 
```{r fig.height = 15, fig.width = 15}
ggraph(ghrg,layout="lgl")+
  geom_edge_link(width=0.2,colour="grey")+
  geom_node_point(col="black",size=0.3)+
  theme_graph()
```

We observe a pattern containg single edge.
We now plot a Dendrogram to show the layout and the extent of the network.

```{r fig.height = 70, fig.width = 70}
ggraph(ghrg, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```



we now extract the edge list from the graph. Then we make a random list that contains indices of 95 percent of the edge list. So we have essentially deleted 5 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 95 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(karate))
random_list <- sample(0.95*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
head(diff,5)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r fig.height = 40, fig.width = 70}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}
pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1
```

##YEAST DATASET

We begin with making a network for the Yeast.We fit the yeast dataset to the Heirarchical Random Graph.
```{r}
data(yeast)
ghrg <- fit_hrg(yeast)
```

We plot the Yeast Dataset on a Graph and check the layout. 
```{r fig.height = 15, fig.width = 15}
ggraph(ghrg,layout="lgl")+
  geom_edge_link(width=0.2,colour="grey")+
  geom_node_point(col="black",size=0.3)+
  theme_graph()
```

We observe a complex pattern which is also entangled with multiple loops along with few networks containg single edge.
We now plot a Dendrogram to show the layout and the extent of the network.

```{r fig.height = 190, fig.width = 70}
ggraph(ghrg, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

By zooming in we can make out the approximate structure of the graph. But it is not very intuitive due to the size and the complexity of the network.

we now extract the edge list from the graph. Then we make a random list that contains indices of 95 percent of the edge list. So we have essentially deleted 5 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 95 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(yeast))
random_list <- sample(0.95*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}

pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1
```



##15 PERCENT KARATE DATASET

We begin with making a network for the karate.We fit the karate dataset to the Heirarchical Random Graph.
```{r}
data(karate)
ghrg <- fit_hrg(karate)
```

we now extract the edge list from the graph. Then we make a random list that contains indices of 85 percent of the edge list. So we have essentially deleted 15 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 85 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(karate))
random_list <- sample(0.85*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
head(diff,5)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r fig.height = 40, fig.width = 70}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}

pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1
```

##15 PERCENT YEAST DATASET

We begin with making a network for the Yeast.We fit the yeast dataset to the Heirarchical Random Graph.
```{r}
data(yeast)
ghrg <- fit_hrg(yeast)
```

We plot the Yeast Dataset on a Graph and check the layout. 
```{r fig.height = 15, fig.width = 15}
ggraph(ghrg,layout="lgl")+
  geom_edge_link(width=0.2,colour="grey")+
  geom_node_point(col="black",size=0.3)+
  theme_graph()
```

We observe a complex pattern which is also entangled with multiple loops along with few networks containg single edge.
We now plot a Dendrogram to show the layout and the extent of the network.

```{r fig.height = 190, fig.width = 70}
ggraph(ghrg, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

By zooming in we can make out the approximate structure of the graph. But it is not very intuitive due to the size and the complexity of the network

we now extract the edge list from the graph. Then we make a random list that contains indices of 85 percent of the edge list. So we have essentially deleted 5 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 85 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(yeast))
random_list <- sample(0.85*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}

pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1 
```


##40 PERCENT KARATE DATASET

We begin with making a network for the karate.We fit the karate dataset to the Heirarchical Random Graph.
```{r}
data(karate)
ghrg <- fit_hrg(karate)
```

we now extract the edge list from the graph. Then we make a random list that contains indices of 60 percent of the edge list. So we have essentially deleted 40 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 60 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(karate))
random_list <- sample(0.60*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
head(diff,5)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r fig.height = 40, fig.width = 70}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}

pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1
```


##40 PERCENT YEAST DATASET

We begin with making a network for the Yeast.We fit the yeast dataset to the Heirarchical Random Graph.
```{r}
data(yeast)
ghrg <- fit_hrg(yeast)
```

We plot the Yeast Dataset on a Graph and check the layout. 
```{r fig.height = 15, fig.width = 15}
ggraph(ghrg,layout="lgl")+
  geom_edge_link(width=0.2,colour="grey")+
  geom_node_point(col="black",size=0.3)+
  theme_graph()
```

We observe a complex pattern which is also entangled with multiple loops along with few networks containg single edge.
We now plot a Dendrogram to show the layout and the extent of the network.

```{r fig.height = 190, fig.width = 70}
ggraph(ghrg, layout = 'dendrogram', circular = FALSE) + 
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

By zooming in we can make out the approximate structure of the graph. But it is not very intuitive due to the size and the complexity of the network

we now extract the edge list from the graph. Then we make a random list that contains indices of 60 percent of the edge list. So we have essentially deleted 5 percent of the edge list. Now we make a new list of the edges that conatins the edgelist of the 40 percent edges. We keep a track of the edges that are deleted in the variable diff which contains the list of edges that were deleted. 

```{r}
d1 <- data.frame(as_edgelist(yeast))
random_list <- sample(0.60*nrow(d1),)
new_list <- data.frame(d1[random_list,])

rows.in.a1.that.are.not.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.without.a2.rows <- a1[!a1.vec %in% a2.vec,]
    return(a1.without.a2.rows)
}

diff <- rows.in.a1.that.are.not.in.a2(d1,new_list)
```

Now we convert the new_list into a matrix then we make a new graph from this edge list. We now fit this graph to a heirarchical random graph.

```{r}
g <- as.matrix(new_list)
g2 <- graph_from_edgelist(g)
ghrg1 <- fit_hrg(g2)
```

We now plot this new heirarchical graph.

```{r}
ggraph(ghrg1, layout = 'dendrogram', circular = FALSE) +
  geom_edge_link() +
  geom_node_point(color="#69b3a2", size=3) +
  geom_node_text(aes( label=name, filter=leaf), angle=0, hjust=1, nudge_y=0.1) +
  geom_node_point(aes(filter=leaf) , alpha=0.6) + ylim(-.6, NA) +
  theme(legend.position="none")+
  coord_flip() +
  scale_y_reverse()
```

We now predict the missing edges in the graph g2. Now we compare the effect of predict function by calculating the difference in the predicted edges and the diff list. 

```{r}

pred <- predict_edges(g2)

rows.in.a1.that.are.in.a2  <- function(a1,a2)
{
    a1.vec <- apply(a1, 1, paste, collapse = "")
    a2.vec <- apply(a2, 1, paste, collapse = "")
    a1.with.a2.rows <- a1[a1.vec %in% a2.vec,]
    return(a1.with.a2.rows)
}

diff2 <- rows.in.a1.that.are.in.a2(diff,pred$edges)
```

Now the difference in the length of the list gives us the accuracy of the prediction.

```{r}
acc1 <- 100*((nrow(diff)-nrow(diff2))/nrow(diff)) 
acc1 
```



