---
title: "ashishsa_hw_p1"
output:
  pdf_document: default
  html_document: default
---
We initially consider the following Utility Matrix
```{r}
library(knitr)
a <- c(4,NA,2)
b <- c(5,3,NA)
c <- c(NA,4,1)
d <- c(5,3,3)
e <- c(1,1,NA)
f <- c(NA,2,4)
g <- c(3,1,5)
h <- c(2,NA,3)
df <- data.frame(a,b,c,d,e,f,g,h)
```
Here we represent the missing values with NA.
```{r}
df
```
We now rename the rows in R.
```{r}
rownames(df) <- c('A','B','C')
df
```
We now need to calculate the Jacard and cosine distance between users by considering the utility matrix as Boolean.
So Initially we convert the utility matrix into Boolean
```{r}
df[df>0] <- 1
df[df<=0] <- 0
df[is.na(df)] <- 0
df
```

We have now converted the Utility Matrix into Binary Utility Matrix. Now we calculate the Jackard Distance.

##JACCARD DISTANCE

This can Manually be calculated as:

The Jaccard Index (between any two columns/users of the matrix M) is a/a+b+c, where:

a = number of rows where both columns are 1

b = number of rows where the first row is 0 and the other row is 1

c = number of rows where the first row is 1 and the other row is 0

Now we calculate the Jaccard distance between (A,B) and (A,C):

Jaccard Distance between A,B:

a=4

b=2

c=2

J_AB=[4/(4+2+2)]

J_AB=[1/2]


Jaccard Distance between A,C:

a=4

b=2

c=2

J_AC=[4/(4+2+2)]

J_AC=[1/2]

##COSINE DISTANCE

COSINE DISTANCE=1-COSINE SIMILARITY

COSINE SIMILARITY=(A.B)/||A||.||B||

##Cosine Distance between A,B:

A.B= (1*0)+(1*1)+(0*1)+(1*1)+(1*1)+(0*1)+(1*1)+(1*0)

A.B=(0+1+0+1+1+0+1+0)

A.B=4

||A||=(1*1)+(1*1)+(0*0)+(1*1)+(1*1)+(0*0)+(1*1)+(1*1)

||A||=6

||B||=(0*0)+(1*1)+(1*1)+(1*1)+(1*1)+(1*1)+(1*1)+(0*0)

||B||=6

COSINE SIMILARITY=4/(6*6)

COSINE SIMILARITY=4/36

COSINE DISTANCE=1-(4/36)

COSINE DISTANCE==32/36

##Cosine Distance between A,C:

A.C= (1*1)+(1*0)+(0*1)+(1*1)+(1*0)+(0*1)+(1*1)+(1*1)

A.C=(1+0+0+1+0+0+1+1)

A.C=4

||A||=(1*1)+(1*1)+(0*0)+(1*1)+(1*1)+(0*0)+(1*1)+(1*1)

||A||=6

||C||=(1*1)+(0*0)+(1*1)+(1*1)+(0*0)+(1*1)+(1*1)+(1*1)

||C||=6

COSINE SIMILARITY=4/(6*6)

COSINE SIMILARITY=4/36

COSINE DISTANCE=1-(4/36)

COSINE DISTANCE==32/36


##Now we consider that 3,4,5 are 1 and 1,2 are 0:
```{r}
library(knitr)
a <- c(4,NA,2)
b <- c(5,3,NA)
c <- c(NA,4,1)
d <- c(5,3,3)
e <- c(1,1,NA)
f <- c(NA,2,4)
g <- c(3,1,5)
h <- c(2,NA,3)
df <- data.frame(a,b,c,d,e,f,g,h)
df[df<=2] <- 0
df[df>=3] <- 1
df[is.na(df)] <- 0
df
```

We have now converted the Utility Matrix into Binary Utility Matrix. Now we calculate the Jackard Distance.

##JACCARD DISTANCE

This can Manually be calculated as:
The Jaccard Index (between any two columns/users of the matrix M) is a/a+b+c, where:

a = number of rows where both columns are 1

b = number of rows where the first row is 0 and the other row is 1

c = number of rows where the first row is 1 and the other row is 0

Now we calculate the Jaccard distance between (A,B) and (A,C):

Jaccard Distance between A,B:

a=2

b=1

c=2

J_AB=[2/(2+1+2)]

J_AB=[2/5]


Jaccard Distance between A,C:

a=2

b=2

c=2

J_AC=[2/(2+2+2)]

J_AC=[2/6]
##COSINE DISTANCE

COSINE DISTANCE=1-COSINE SIMILARITY

COSINE SIMILARITY=(A.B)/||A||.||B||

##Cosine Distance between A,B:
A.B= (1*0)+(1*1)+(0*1)+(1*1)+(0*0)+(0*0)+(1*0)+(0*0)

A.B=(0+1+0+1+0+0+0+0)

A.B=2

||A||=(1*1)+(1*1)+(0*0)+(1*1)+(0*0)+(0*0)+(1*1)+(0*0)

||A||=4

||B||=(0*0)+(1*1)+(1*1)+(1*1)+(0*0)+(0*0)+(0*0)+(0*0)

||B||=3

COSINE SIMILARITY=2/(4*3)

COSINE SIMILARITY=2/12

COSINE DISTANCE=1-(2/12)

COSINE DISTANCE==10/12

##Cosine Distance between A,C:
A.C= (1*0)+(1*0)+(0*0)+(1*1)+(0*0)+(0*1)+(1*1)+(0*1)

A.C=(0+0+0+1+0+0+1+0)

A.C=2

||A||=(1*1)+(1*1)+(0*0)+(1*1)+(0*0)+(0*0)+(1*1)+(0*0)

||A||=4

||C||=(0*0)+(0*0)+(0*0)+(1*1)+(0*0)+(1*1)+(1*1)+(1*1)

||C||=4

COSINE SIMILARITY=2/(4*4)

COSINE SIMILARITY=2/16

COSINE DISTANCE=1-(2/16)

COSINE DISTANCE==14/16

Normalize the matrix by subtracting from each nonblank entry the average value
for its user. Using this matrix, compute the cosine distance between each pair of
users. 

We initially consider the following Utility Matrix
```{r}
library(knitr)
a <- c(4,NA,2)
b <- c(5,3,NA)
c <- c(NA,4,1)
d <- c(5,3,3)
e <- c(1,1,NA)
f <- c(NA,2,4)
g <- c(3,1,5)
h <- c(2,NA,3)
df <- data.frame(a,b,c,d,e,f,g,h)
```

```{r}
A <- c(4,5,NA,5,1,NA,3,2)
B <- c(NA,3,4,3,1,2,1,NA)
C <- c(2,NA,1,3,NA,4,5,3)

mean_A <- mean(A,na.rm = TRUE)
mean_A
```


```{r}
mean_B <- mean(B,na.rm = TRUE)
mean_B
```


```{r}
mean_C <- mean(C,na.rm = TRUE)
mean_C
```

Now substracting mean from every value in df we get:
```{r}
A <- c(1,2,NA,2,-2,NA,0,-1)
B <- c(NA,1,2,1,-1,0,-1,NA)
C <- c(-1,NA,-2,0,NA,1,2,0)
```

##COSINE DISTANCE

COSINE DISTANCE=1-COSINE SIMILARITY

COSINE SIMILARITY=(A.B)/||A||.||B||

we consider na as 0
##Cosine Distance between A,B:

A.B= (1*0)+(2*1)+(0*2)+(2*1)+(-2*-1)+(0*0)+(0*-1)+(-1*0)

A.B=(0+2+0+2+2+0+0+0)

A.B=6

||A||=(1*1)+(2*2)+(0*0)+(2*2)+(-2*-2)+(0*0)+(0*0)+(-1*-1)

||A||=14

||B||=(0*0)+(1*1)+(2*2)+(1*1)+(-1*-1)+(0*0)+(-1*-1)+(0*0)

||B||=8

COSINE SIMILARITY=6/(14*8)

COSINE SIMILARITY=2/112

COSINE DISTANCE=1-(2/112)

COSINE DISTANCE==110/112

##Cosine Distance between A,C:
A.C= (1*0)+(1*0)+(0*0)+(1*1)+(0*0)+(0*1)+(1*1)+(0*1)

A.C=(0+0+0+1+0+0+1+0)

A.C=2

||A||=(1*1)+(2*2)+(0*0)+(2*2)+(-2*-2)+(0*0)+(0*0)+(-1*-1)

||A||=14


||C||=(-1*-1)+(0*0)+(-2*-2)+(0*0)+(0*0)+(1*1)+(2*2)+(0*0)

||C||=10

COSINE SIMILARITY=2/(14*10)

COSINE SIMILARITY=2/140

COSINE DISTANCE=1-(2/140)

COSINE DISTANCE==138/140

