---
title: "ashishsa_hw3_p3"
output:
  pdf_document: default
  html_document: default
---

We perform PCA on 200 bank notes which contains 100 geniuine and 100 counterfeit bank notes.
```{r warning=FALSE,results=FALSE,comment=FALSE}
rm(list = ls())
library(pca3d)
library(purrr)
library(kohonen)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(factoextra)
library(gridExtra)
library(cowplot)
library(cluster)
library(lfda)
load("SwissBankNotes.RData")
df1 <- SwissBankNotes
df1 <- as.data.frame(df1)
```

We know that the first 100 Bank Notes are Real and the Second 100 Bank Notes are Counterfeit. We now add one more column to the original dataframe which states that first 100 bank notes are True and second 100 bank notes are False.

```{r}
status <- c(rep('TRUE',len=100),rep('FALSE',len=100))
df2 <- cbind(df1,status)
```

We now convert the last column into factors.

```{r}
df2$status <- as.factor(as.character(df2$status))
df2$status <- df2$status %>% map_if(is.factor, as.numeric)
df2$status <- as.numeric(df2$status)
```


```{r fig.width=30, fig.height=15}
plot1 <- ggplot(df1,aes(x=factor(0),df1$length))+geom_boxplot()+
   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot2 <- ggplot(df1,aes(x=factor(0),df1$height.left))+geom_boxplot()+
   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot3 <- ggplot(df1,aes(x=factor(0),df1$height.right))+geom_boxplot()+theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot4 <- ggplot(df1,aes(x=factor(0),df1$inner.lower))+geom_boxplot()+
   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot5 <- ggplot(df1,aes(x=factor(0),df1$inner.upper))+geom_boxplot()+
   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot6 <- ggplot(df1,aes(x=factor(0),df1$diagonal))+geom_boxplot()+
   theme(axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank())
plot_grid(plot1, plot2, plot3,plot4,plot5,plot6,ncol=6)
```

We first begin by performing Boxplot Analysis. We first analyze length and observe that it has outliers at 213,215.7 and 217.
We analyze height.left and observe that it has outlier at 129.
We analyze height.right and observe that it has outliers at 131.5.
We analyze inner.lower and observe that it has no outliers.
We analyze inner.upper and observe that it has outliers at 8.
We analyze diagonal and observe that it has no outliers.

From the Boxplot Analysis we conclude that the Variables need to be scaled as inner upper and inner lower are of different dimension in comparison to the other variables.

```{r}
m <- apply(df1,2,mean)
s <- apply(df1,2,sd)
z <- scale(df1,m,s)
```

We first Analyze all the variables that is fake and real notes together.
We now apply Principal Component Analysis on dataset in R.

```{r}
z_pca <- prcomp(z,center = TRUE,scale. = FALSE)
```

We now plot the Eigen Values for the entire dataset.

```{r}
fviz_eig(z_pca)
```

We summarize the Results for the PCA and observe here that PC1 only accounts for 49Percent of the Variation. While PC2 only accounts for 21Percent of the Variation. Since we have already used scaled data for plotting the dimension of the variables has no significant impact.
This indicates that the data that is skewed by is because of false data. 

```{r}
summary(z_pca)
```

We now check the Eigen Values for each of the 6 variables.

```{r}
EigenValues <- eigen(cov(scale(df1)))
EigenValues$values
```

The Eigen Values obtained above are checked using the following method. We calculate the Standard Deviation of the Principle Components and find the square of it. The Values must be same indicating that they are the intended correct value. 

```{r}
(z_pca$sdev)^2
```

We first plot a score plot. The score plot plots each of the bank notes (Real or Fake) against each of the Principle Components.

```{r fig.width=30, fig.height=15}
PCAcolors <- c("#66c2a5","#8da0cb")[unique(as.integer(df2$status))]


PCAscores <- z_pca$x
PCAloadings <- z_pca$rotation


par(mfrow=c(1,2))
{plot(PCAscores,  
     pch=21,           
     col=PCAcolors,    
     bg=PCAcolors,     
     cex=1.5,          
     main="Scores"     
)
legend("topright",
       legend=unique(df2$status),                       
       pch=21,                                    
       pt.bg=c("#66c2a5","#8da0cb"),
       pt.cex=1.5,                               
       col = c("#66c2a5","#8da0cb"))
}

{plot(PCAloadings[,1:2],   
     pch=21,              
     bg="black",          
     cex=1,              
     main="Loadings"     
)
text(PCAloadings[,1:2],             
     labels=rownames(PCAloadings)  
)}
```

From the score plot we can see a clear distinction between the Real and the Fake Notes. The Real Notes have principle components towards the lower half of the plot.

```{r fig.width=30, fig.height=15}
autoplot(fanny(df2[,-7],2),frame=TRUE)
```

We now map the positively and negatively correlated components for all the True and Fake Notes. We observe the components for each of the fake and real notes and see the 2 clusters (True and Fake Notes)and the components affecting it positively and negatively.

```{r fig.width=30, fig.height=15}
fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = FALSE)
```

We now map the positively and negatively correlated components for all the True and Fake Notes. We observe the components and the corresponding vectors for each of the fake and real notes and see the 2 clusters (True and Fake Notes)and the components affecting it positively and negatively and its spread across the feature space.

```{r fig.width=30, fig.height=15}
fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE)
```

We now map the positively and negatively correlated components for all the True and Fake Notes in the first and third principal components. We observe the components for each of the fake and real notes and see the 2 clusters (True and Fake Notes)and the components affecting it positively and negatively.

```{r fig.width=30, fig.height=15}
fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = FALSE,axes = c(1, 3))
```

We now map the positively and negatively correlated components for all the True and Fake Notes in the first and third principal components. We observe the components and the corresponding vectors for each of the fake and real notes and see the 2 clusters (True and Fake Notes)and the components affecting it positively and negatively and its spread across the feature space.

```{r fig.width=30, fig.height=15}
fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
```

We now plot the Eigen Vectors across the sample space and observe the distribution of the factors across each of the principal components and study the effect of each of these factors across the sample space.

```{r fig.width=30, fig.height=15}
fviz_pca_var(z_pca,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE
)
```

We can conclude that height(right,left), diagonal, length has the highest contribution across PC1 and PC2 while inner upper has the lowest contribution. inner lower has a moderate contribution to the principle space.

We now plot the Eigen Vectors across the sample space and observe the distribution of the factors across each of the principal components 1 and 3 and study the effect of each of these factors across the sample space.

```{r fig.width=30, fig.height=15}
fviz_pca_var(z_pca,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE,
axes = c(1, 3)
)
```

We can conclude that height(right,left), diagonal, inner lower has a moderate contribution to the principle space.length has the lowest contribution to the principle space.inner upper has the highest contribution across PC1 and PC3 while inner upper has the highest contribution.

We now find the principle components for the Real and Fake Notes Separately.

```{r}
d1 <- scale(df2[1:100,-7])
d2 <- scale(df2[101:200,-7])
z_pca1 <- prcomp(d1,center = TRUE,scale. = FALSE)
z_pca2 <- prcomp(d2,center = TRUE,scale. = FALSE)
```

We now plot the Eigen vectors for 
1)Both True and False Notes
2)True Notes
3)False Notes

```{r fig.width=30, fig.height=15}
plot1 <- fviz_eig(z_pca)
plot2 <- fviz_eig(z_pca1)
plot3 <- fviz_eig(z_pca2)
plot_grid(plot1, plot2, plot3,ncol = 3)
```

The above plot shows comparison of the Percentage Of Variance for 1)Both Real and Fake 2) True 3)Fake notes. We observe that for Real Notes we have highest variances across PC1,PC2. We observe that for Fake Notes we have highest variance across PC1 while all other components are scarcely distributed.

The first plot indicates the Eigen values for True Bank Notes.

```{r fig.width=30, fig.height=15}
plot(plot2)
```

We observe from the first plot that PC1 accounts for 49 Percent variance while PC2 accounts for 28 Percent Variance.

The second plot indicates the Eigen values for Fake Bank Notes.

```{r fig.width=30, fig.height=15}
plot(plot3)
```

We observe from the second plot that PC1 accounts for 32 Percent variance while PC2 accounts for 29 Percent Variance.

We now summarize the PCA for both True and Fake Bank Notes.

```{r}
summary(z_pca1)
summary(z_pca2)
```

We find the Eigen Values for the first 100 values ie., The True Bank Notes.

```{r}
EigenValues1 <- eigen(cov(d1))
EigenValues1$values
```

The Eigen Values obtained above are checked using the following method. We calculate the Standard Deviation of the Principle Components and find the square of it. The Values must be same indicating that they are the intended correct value.

```{r}
(z_pca1$sdev)^2
```

We find the Eigen Values for the second 100 values ie., The Fake Bank Notes.

```{r}
EigenValues2 <- eigen(cov(d2))
EigenValues2$values
```

The Eigen Values obtained above are checked using the following method. We calculate the Standard Deviation of the Principle Components and find the square of it. The Values must be same indicating that they are the intended correct value. 

```{r}
(z_pca2$sdev)^2
```

We now plot the principle components and then plot the variables across the principle components. plot 2 indicates the True Bank Notes while the plot3 indicates the False Bank Notes.

```{r fig.width=30, fig.height=15}
plot1 <- fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE)
plot2 <- fviz_pca_ind(z_pca1,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE)
plot3 <- fviz_pca_ind(z_pca2,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE)
plot_grid(plot1, plot2, plot3,ncol=3)
```

We now plot the Individual PCA components and also the Variables that affect the PCA. We observe that the True Bank Notes have higeher contribution from Inner(lower,upper), height(left,right), medium contribution from length,low contribution from diagonal.

```{r fig.width=30, fig.height=15}
plot2
```


```{r fig.width=30, fig.height=15}
plot2
plot2_1 <- fviz_pca_var(z_pca1,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE
)
plot_grid(plot2, plot2_1,ncol = 2)
```

```{r fig.width=30, fig.height=15}
plot3
```

We now plot the Individual PCA components and also the Variables that affect the PCA. We observe that the False Bank Notes have higeher contribution from Inner(lower) and height(left,right), Inner(upper) has high contribution.length has medium contribution.lowest contribution from diagonal.

```{r fig.width=30, fig.height=15}
plot3
plot3_1 <- fviz_pca_var(z_pca2,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE
)
plot_grid(plot3,plot3_1,ncol = 2)
```

We now plot the Principle Components across First and Third Dimension.

```{r fig.width=30, fig.height=15}
plot1 <- fviz_pca_ind(z_pca,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
plot2 <- fviz_pca_ind(z_pca1,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
plot3 <- fviz_pca_ind(z_pca2,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
plot_grid(plot1, plot2, plot3,ncol=3)
```

```{r fig.width=30, fig.height=15}
plot2
```


We now plot the Individual PCA components across dimensions 1,3 and also the Variables that affect the PCA. We observe that the True Bank Notes have higher contribution from Inner(lower,upper), height(left,right), medium contribution from length,low contribution from diagonal.

```{r fig.width=30, fig.height=15}
plot2 <- fviz_pca_ind(z_pca1,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
plot2_1 <- fviz_pca_var(z_pca1,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE,
axes = c(1, 3)
)
plot_grid(plot2, plot2_1,ncol = 2)
```

```{r fig.width=30, fig.height=15}
plot3
```

We now plot the Individual PCA components across dimensions 1,3 and also the Variables that affect the PCA. We observe that the False Bank Notes have higher contribution from diagonal.Inner(lower,upper), height(left) has medium contribution.length, weight right has low contribution.


```{r fig.width=30, fig.height=15}
plot3 <- fviz_pca_ind(z_pca2,col.ind = "cos2",gradient.cols=c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE,axes = c(1, 3))
plot3_1 <- fviz_pca_var(z_pca2,
col.var = "contrib",
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE,
axes = c(1, 3)
)
plot_grid(plot3, plot3_1,ncol = 2)
```




